\section*{Parameter governance}
\label{sec:parameter-governance}

\paragraph{Units and scaling.}
We use the unit system defined in \S\ref{sec:strong-forms-nd}.

\paragraph{Mechanical parameters.}
Apparent densities in human cancellous bone span roughly $0.05$--$1.1~\mathrm{g/cm^3}$, while cortical bone is commonly reported around $1.7$--$2.0~\mathrm{g/cm^3}$ \cite{Helgason2008,Rho1993}.
This motivates bounds $\rho_{\min}$ (numerical floor) and $\rho_{\max}$; in practice $\rho_{\min}$ primarily serves as a conditioning parameter for the mechanics.
For the isotropic baseline, $\nu_0\approx 0.3$ is typical and cortical Young's moduli are on the order of $15$--$25~\mathrm{GPa}$ \cite{Rho1993}.
For the density--stiffness scaling in \eqref{eq:zysset-curnier}, adaptive remodeling studies commonly use power-law exponents in the range $n_{\mathrm{trab}}\approx 2$--$3$ for trabecular bone and $n_{\mathrm{cort}}\approx 1$--$2$ for denser tissue \cite{Weinans1992_BehaviorAdaptiveBoneRemodeling,Keller1994_PredictingCompressiveMechanicalBehavior,Helgason2008}.
The transition densities $\rho_{\mathrm{trab}}^{\max}$ and $\rho_{\mathrm{cort}}^{\min}$ should bracket the trabecular--cortical overlap so that the blended exponent $k(\widetilde{\rho})$ varies only over a physiologically plausible interval.
Finally, the fabric stiffness exponents $p_E$ and $p_G$ control how strongly anisotropy is expressed in stiffness; values near $1$ are typical in fabric-based orthotropic scaling \cite{Zysset1995_AlternativeModelAnisotropic}, and $p_E,p_G\in[0.5,2]$ is a conservative prior unless a dedicated calibration dataset is available.

\paragraph{Stimulus parameters.}
The mechanostat is driven by the \emph{specific} strain-energy density $m=\psi/\rho$; in this unit system $m$ is naturally reported in J/g (numerically equal to MPa divided by g/cm$^3$).
Peak physiological strains in load-bearing bone are on the order of hundreds to a few thousand microstrain \cite{Turner1998_ThreeRulesBone,Robling2006}.
For apparent moduli in the $1$--$20~\mathrm{GPa}$ range this implies strain-energy densities on the order of $10^{-3}$--$10^{-1}~\mathrm{MPa}$ (via $\psi\sim\tfrac12 E\varepsilon^2$); corresponding specific energies $m=\psi/\rho$ are therefore load-dependent and should be calibrated once the loading magnitude and stiffness scaling are fixed.
Weinans et al.\ underscored that the reference values $(\psi_{\mathrm{ref}}^{\mathrm{trab}},\psi_{\mathrm{ref}}^{\mathrm{cort}})$ are ultimately calibration parameters tied to the applied loads.
The averaging exponent $p$ in \eqref{eq:stim-psi-nd} controls how strongly $\psi_{\mathrm{day}}$ is peak-biased; $p\in[1,6]$ spans the range from a weighted mean ($p=1$) to a strongly peak-dominated driver.
The lazy-zone half-width $\delta_0$ controls the tolerated deviation around the set-point before net formation/resorption; practical values are typically $\delta_0\approx 0.05$--$0.3$.
The saturation parameters $(S_{\max},\kappa_S)$ should bound $S$ without clipping physiologically relevant values; $S_{\max}\sim 0.5$--$2$ with $\kappa_S\sim 0.1$--$1$ are robust defaults.
For the transport parameters, $\tau_S$ should be short relative to density remodeling (days to a few weeks), and $D_S$ can be chosen to produce a sensing length $\ell_S=\sqrt{\tau_S D_S}$ on the order of $10^{-1}$--$10^{0}~\mathrm{mm}$, depending on the intended degree of nonlocal averaging.

\paragraph{Density parameters.}
Weinans et al.\ reported $\rho_{\max}\approx 1.74~\mathrm{g/cm^3}$ and a near-resorption floor $\rho_{\min}\approx 0.01~\mathrm{g/cm^3}$ in simulation studies \cite{Weinans1992_BehaviorAdaptiveBoneRemodeling}.
In practice, larger numerical floors (e.g.\ $\rho_{\min}\gtrsim 0.05$) improve conditioning if very low-density regions cause mechanical ill-conditioning, while keeping $\rho_{\min}$ small is appropriate if complete resorption is part of the intended biology.
The kinetic gains $(k_{\mathrm{form}},k_{\mathrm{res}})$ have units of density per time and set the dominant remodeling time scale; a convenient summary quantity is the characteristic reaction rate $r_{\mathrm{char}}=\max(k_{\mathrm{form}}/\rho_{\max},\,k_{\mathrm{res}}/\rho_{\min})$.
To represent adult adaptation on the order of months, implied characteristic rates are typically $O(10^{-3}$--$10^{-2})~\mathrm{d^{-1}}$ when $S=O(1)$ \cite{Robling2006}.
The diffusion coefficient $D_\rho$ primarily acts as a regularization length scale; choosing an effective diffusion length on the order of $0.5$--$2~\mathrm{mm}$ (or a few element sizes) is generally sufficient to suppress mesh-scale oscillations without over-smoothing (see the mesh-based prescription below).
If the surface-availability factor $\alpha_{\mathrm{surf}}(\rho)$ is enabled (cf.\ \eqref{eq:Asurf-nd}), the minimum activity $\alpha_{\min}$ and the half-saturation scale $S_0$ should prevent kinetics from freezing at high density while still reflecting reduced available surface in dense tissue \cite{Pivonka2013_InfluenceBoneSurfaceAvailability}.

\paragraph{Fabric parameters.}
Microstructural reorientation is expected to evolve on the remodeling time scale (months) rather than on the loading time scale; consequently $\tau_A\sim 30$--$300~\mathrm{d}$ is a reasonable prior if fabric represents remodeling-driven alignment \cite{Robling2006}.
The diffusion coefficient $D_A$ controls the spatial smoothing of $\mathbf{L}$ and can be tuned to match the targeted correlation length of anisotropy.
The exponent $\gamma_F$ controls how strongly stress-eigenvalue contrasts translate into fabric-eigenvalue ratios; $\gamma_F\in[0.5,2]$ is conservative (with $\gamma_F=1$ a neutral default).
The bounds $m_{\min},m_{\max}$ cap admissible anisotropy; the ratio $m_{\max}/m_{\min}$ limits the maximum eigenvalue ratio and should be chosen to avoid numerically extreme stiffness contrasts.
Finally, the regularization parameters $\varepsilon_Q$ and $\varepsilon_a$ should be small relative to the scale of $\bar{\mathbf{Q}}$ so that they only stabilize eigen-decomposition and activity gating near isotropy.

\subsection*{Parameter sets for validation and benchmarking}
To assess both physiological plausibility and numerical robustness, we use two parameter configurations for the box benchmarks: the \textit{PhysioSetup} (baseline) and the \textit{StiffSetup} (stress test), summarized in Table~\ref{tab:param_setup}.
The \textbf{StiffSetup} narrows the mechanostat ($\delta_0,\kappa_S$), accelerates density kinetics ($k_{\mathrm{form}},k_{\mathrm{res}}$), and lowers the density floor $\rho_{\min}$, which induces sharper internal gradients and faster transients that challenge the convergence of fixed-point coupling.

\begin{table*}[!htbp]
  \centering
  \footnotesize
  \setlength{\tabcolsep}{4pt}
  \renewcommand{\arraystretch}{1.15}
  \caption{Physical parameter sets used in the box numerical experiments: \textbf{PhysioSetup} (baseline) and \textbf{StiffSetup} (stress test).}
  \label{tab:param_setup}
  \begin{tabularx}{\textwidth}{@{} >{\raggedright\arraybackslash}X l l r r @{}}
    \toprule
    \textbf{Parameter} & \textbf{Symbol} & \textbf{Unit} & \textbf{PhysioSetup} & \textbf{StiffSetup} \\
    & & & \textit{(Baseline)} & \textit{(Hard)} \\
    \midrule
    \multicolumn{5}{c}{\textbf{I. Setup-dependent parameters}} \\
    \midrule
    \textit{Stimulus \& mechanostat} & & & & \\
    Lazy-zone half-width & $\delta_0$ & -- & $0.20$ & $0.10$ \\
    Saturation width & $\kappa_S$ & -- & $0.50$ & $0.30$ \\
    Stimulus diffusivity & $D_S$ & mm$^2$/day & $0.225$ & $0.0225$ \\
    \addlinespace
    \textit{Density evolution} & & & & \\
    Density floor & $\rho_{\min}$ & g/cm$^3$ & $0.20$ & $0.10$ \\
    Formation gain & $k_{\mathrm{form}}$ & g/cm$^3$/day & $0.01$ & $0.05$ \\
    Resorption gain & $k_{\mathrm{res}}$ & g/cm$^3$/day & $0.0025$ & $0.01$ \\
    Density diffusivity & $D_\rho$ & mm$^2$/day & $0.2025$ & $0.02025$ \\
    \midrule
    \multicolumn{5}{c}{\textbf{II. Shared parameters}} \\
    \midrule
    \textit{Elasticity (Zysset--Curnier)} & & & & \\
    Reference Young's modulus & $E_0$ & MPa & \multicolumn{2}{c}{$7500$} \\
    Poisson's ratio & $\nu_0$ & -- & \multicolumn{2}{c}{$0.3$} \\
    Stiffness exponents (trab./cort.) & $n_{\mathrm{trab}}, n_{\mathrm{cort}}$ & -- & \multicolumn{2}{c}{$2.0, \ 1.3$} \\
    Transition densities (trab./cort.) & $\rho_{\mathrm{trab}}^{\max}, \rho_{\mathrm{cort}}^{\min}$ & g/cm$^3$ & \multicolumn{2}{c}{$1.0, \ 1.25$} \\
    Fabric stiffness exponents & $p_E, p_G$ & -- & \multicolumn{2}{c}{$1.0, \ 1.0$} \\
    \addlinespace
    \textit{Stimulus transport \& reference} & & & & \\
    SED averaging exponent & $p$ & -- & \multicolumn{2}{c}{$4.0$} \\
    Reference specific SED (trab.) & $\psi_{\mathrm{ref}}^{\mathrm{trab}}$ & J/g & \multicolumn{2}{c}{$5\times 10^{-5}$} \\
    Reference specific SED (cort.) & $\psi_{\mathrm{ref}}^{\mathrm{cort}}$ & J/g & \multicolumn{2}{c}{$5\times 10^{-5}$} \\
    Time constant & $\tau_S$ & day & \multicolumn{2}{c}{$10$} \\
    Max stimulus magnitude & $S_{\max}$ & -- & \multicolumn{2}{c}{$1.0$} \\
    \addlinespace
    \textit{Density bounds \& surface availability} & & & & \\
    Initial density & $\rho_0$ & g/cm$^3$ & \multicolumn{2}{c}{$1.0$} \\
    Upper density bound & $\rho_{\max}$ & g/cm$^3$ & \multicolumn{2}{c}{$2.0$} \\
    Tissue density & $\rho_{t}$ & g/cm$^3$ & \multicolumn{2}{c}{$2.0$} \\
    Surface availability enabled & -- & -- & \multicolumn{2}{c}{yes} \\
    Minimum activity & $\alpha_{\min}$ & -- & \multicolumn{2}{c}{$0.02$} \\
    Half-saturation scale & $S_0$ & 1/mm & \multicolumn{2}{c}{$1.0$} \\
    \addlinespace
    \textit{Fabric evolution} & & & & \\
    Time constant & $\tau_A$ & day & \multicolumn{2}{c}{$120$} \\
    Fabric diffusivity & $D_A$ & mm$^2$/day & \multicolumn{2}{c}{$0.0188$} \\
    Fabric exponent & $\gamma_F$ & -- & \multicolumn{2}{c}{$1.0$} \\
    Fabric regularization & $\varepsilon_Q$ & MPa$^2$ & \multicolumn{2}{c}{$10^{-12}$} \\
    Activity-gate width & $\varepsilon_a$ & -- & \multicolumn{2}{c}{$10^{-4}$} \\
    Anisotropy bounds & $[m_{\min}, m_{\max}]$ & -- & \multicolumn{2}{c}{$[0.2, \ 5.0]$} \\
    \bottomrule
  \end{tabularx}
\end{table*}

\paragraph{Time step size.}
A conservative time-step choice follows the fastest intrinsic time scales: in \textbf{PhysioSetup} the stimulus relaxes over $\tau_S=10~\mathrm{d}$ and fabric over $\tau_A=120~\mathrm{d}$, so resolving stimulus transients suggests $\Delta t \sim 0.05\!-\!0.2\,\tau_S \approx 0.5\!-\!2~\mathrm{d}$; density kinetics are typically slower and therefore less restrictive.
In \textbf{StiffSetup}, accelerated formation/resorption and a smaller density floor shorten the density time scale, so a conservative bound is
\[
\Delta t \sim 0.05\!-\!0.2\min(\rho_{\max}/k_{\mathrm{form}},\,\rho_{\min}/k_{\mathrm{res}}),
\]
which is of the same order as $\tau_S$ for the values in Table~\ref{tab:param_setup}.
Expected settling to quasi-steady behavior occurs over a few time constants: stimulus $\sim 3\tau_S$ (tens of days), fabric $\sim 3\tau_A$ (order $10^2$ days), and density over a few multiples of its dominant kinetic scale (months in Physio, potentially weeks in Stiff).
Note that numerical convergence can still hold for much larger $\Delta t$ because the stimulus is bounded and saturating ($\tanh$, $|S|\le S_{\max}$) and density rates self-limit near $\rho_{\min},\rho_{\max}$; such large steps should be viewed as steady-state accelerators unless verified by step-refinement ($\Delta t$ vs.\ two $\Delta t/2$).

\paragraph{Regularization diffusivities.}
Let $h$ denote a characteristic element size (e.g., local mean edge length, or $h=\sqrt[3]{V_e}$ from the element volume). To obtain mesh-insensitive fields and suppress the ``patchwork'' response of purely local adaptive remodeling rules, we prescribe a target regularization length $\ell=\alpha h$ with $\alpha\in[1,2]$ \cite{Weinans1992_BehaviorAdaptiveBoneRemodeling,Peerlings1998,Giorgio2019_MechanicallyDrivenBiological}. Using the intrinsic lengths $\ell_S=\sqrt{\tau_S D_S}$, $\ell_A=\sqrt{D_A\tau_A/a(\bar{\mathbf{Q}})}$ (steady operator), and $\ell_\rho\approx\sqrt{D_\rho/(\Delta t^{-1}+r_{\mathrm{char}})}$ (implicit density update), we set
\begin{equation}
D_S=\frac{\ell^2}{\tau_S},\qquad
D_A=\frac{\ell^2}{\tau_A}\ \ \text{(use $a(\bar{\mathbf{Q}})=1$ for the minimal length)},\qquad
D_\rho=\ell^2\left(\Delta t^{-1}+r_{\mathrm{char}}\right),
\end{equation}
where the conservative (smallest-length) reaction estimate is $r_{\mathrm{char}}=\max\!\left(k_{\mathrm{form}}/\rho_{\max},\,k_{\mathrm{res}}/\rho_{\min}\right)$ evaluated at maximal activity (e.g., $S_\pm\approx 1$ and $\alpha_{\mathrm{surf}}\approx 1$). In the quasi-static regime $\Delta t^{-1}\ll r_{\mathrm{char}}$ this reduces to $D_\rho\approx \ell^2 r_{\mathrm{char}}$.

\paragraph{Box benchmark.}
For the reference box mesh ($10\times 10\times 30$~mm, $n_x=n_y=10$, $n_z=30$) we take $h\approx 1$~mm and choose $\alpha=1.5$, hence $\ell\approx 1.5$~mm.
With $\tau_S=10$~d and $\tau_A=120$~d, the mesh-based estimates give $D_S=\ell^2/\tau_S\approx 0.225$~mm$^2$/day and $D_A=\ell^2/\tau_A\approx 0.0188$~mm$^2$/day.
Using $\Delta t=25$~d and a conservative $r_{\mathrm{char}}=0.05~\mathrm{d^{-1}}$ yields $D_\rho\approx \ell^2(\Delta t^{-1}+r_{\mathrm{char}})\approx 0.2025$~mm$^2$/day.
In the reference configurations we intentionally under-regularize stimulus and density by a factor $0.1$, i.e.\ $D_S=0.0225$~mm$^2$/day and $D_\rho=0.02025$~mm$^2$/day, while keeping $D_A$ unchanged.

\paragraph{Femur benchmark.}
For the femur mesh, the mean element length is $h\approx 1.36$~mm.
Using $\alpha=1.5$, the target regularization length is $\ell\approx 2.04$~mm.
For the \textbf{StiffSetup} parameters ($r_{\mathrm{char}}=0.05~\mathrm{d^{-1}}$), the baseline estimates are $D_S\approx 0.415$~mm$^2$/day, $D_A\approx 0.0346$~mm$^2$/day, and $D_\rho\approx 0.373$~mm$^2$/day.
Applying the same $0.1$ under-regularization factor for stimulus and density yields the values: $D_S=0.0415$~mm$^2$/day, $D_\rho=0.0373$~mm$^2$/day, and $D_A=0.0346$~mm$^2$/day.
